---
title: "R, logger and working smarter"
author: "Dario Mavec"
date: 2020-07-20
categories: ["R"]
tags: ["logger", "R", "glue"]
---

# The Forest

Imagine with me a moment.

You've spent all day writing killer R code. It has been a blur of data munging, model tweaks and picture perfect analysis outputs, using test data.

The clock hits 6pm, and you set your code running against the full dataset. By your estimate, it should take about 8 hours to run. Perfect! You can start fresh in the morning with a big, strong coffee. Only when you start the next day, you discover that the job has failed. Worse than that, the application crashed so you can't even get any information on what caused the crash.

Uhoh... Time for a day of tinkering away at code, trying to figure out what went wrong. Maybe you make a few fixes and try again that night, only to discover it crashes **again**. At this point, you might as well swap the console for the 9-tails - it can't be any more painful!

How about we work smarter rather than harder.

# **logger** to the rescue! `r emo::ji('evergreen_tree')`

Using the excellent [logger](https://github.com/daroczig/logger) package, you can gather way more information about jobs and simplify the debugging process. Once loaded we immediately have access to a suite of `log_*()` functions to track job progress.

```{r, message=FALSE, warning=FALSE, collapse=TRUE}
if (!require("logger")) install.packages("logger")
library(logger)

log_debug("I'm a sneaky little log")
log_success("I'm a happy little log")
log_info("I'm a friendly little log")
log_warn("I'm a worrying little log")
log_error("I'm a scary little log")
log_fatal("Uhoh! I'm a dead little log")
```
Now you may be asking, why can't you just use `print()` or `message()` like you always have? Well I'm glad you asked! Allow me to demonstrate some of the benefits.

## The Benefits

First, the obvious. Every messages now has a timestamp, which can be incredibly useful for identifying and improving slow code. We also have a bunch of different severity levels for the messages which allow us to be much more nuanced with our logging. I find it useful to use `log_debug()` as I'm developing and just go totally nuts with how many logs I include. By default these messages won't appear, however by changing the log threshold we can see them:

```{r, collapse=TRUE}
log_debug("Nyah, you can't see me!")
log_threshold(DEBUG)
log_debug("Oh noes!")
```

## String Interpolation

Now for some of the niftier features! The log functions come with [glue](https://github.com/tidyverse/glue) strings enabled by default. `glue` adds Python-like string interpolation, which simply means we can refer to variables inside the string using `{}` marks. Anything inside of the brackets will get evaluated and, assuming its printable, added to the string.

```{r, collapse=TRUE}
for (i in 1:3) {
 log_info("i = {i}") 
}
```

Suddenly we have a super easy way to track progress when processing large loops. Another great alternative for monitoring loops is the `progress` package, which Dan has a great intro to over [here](http://gradientdescending.com/fun-with-progress-bars-fish-daggers-and-the-star-wars-trench-run/).

## Log to File

Logging to console is all well and good, but in our scenario we've had a double calamity and our process has completely died. That means no more nice console messages. Fortunately we can simultaneously output to file and console at the same time with the nifty `appender_tee()`.

```{r, eval=FALSE}
log_appender(appender_tee("path/to/my/log_file.log"))
```
```{r, collapse=TRUE}
log_info("Woah, I'm in two places at once...")
```

When a process fails we now have a file we can look at. Also, it's now possible for any member of a team to check on the status of a long running job.

If you want to get **really** fancy, you can even shoot the logs over to a Slack channel using `appender_slack()`. I would recommend only doing this for `SUCCESS` or `ERROR` level messages, otherwise you might end up with hella spam.

# In Summary...

`logger` is awesome, and has saved me countless debugging hours. Check it out!

Mad props to [daroczig](https://github.com/daroczig). And keep on sharpening that sword! `r emo::ji('sword')`
